The Power of Unbiased Recursive Partitioning:\\ A Unifying View of CTree, MOB, and GUIDE

### 2019-03-05 pkg/partykit/inst/guideliketest/DAGStat2019/slides.Rnw
small corrections in notation and illustrations
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2482 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: cca523f738777eba594966e8394465dd42c01660

{\bf General strategy:}
\begin{enumerate}
\item Evaluate a discrepancy measure giving information on the goodness of fit of the model for each observation and each model parameter (transformation~of~$Y$).
\item Apply a statistical test assessing dependency of the discrepancy measure to each possible split variable $Z_j$.
\item Select the split variable $Z_j^\ast$ showing the smallest $p$-value.
\end{enumerate}

@@

{\bf Residuals:}
\begin{flalign*}
\hspace{0.7cm}
r(y,x,\hat{\beta_0},\hat{\beta_1}) = y - \hat{\beta_0} - \hat{\beta_1} \cdot x &&
\end{flalign*}

@@

{\bf Model scores} (based on least squares):
\begin{flalign*}
\hspace{0.7cm}
s(y, x, \hat{\beta_0}, \hat{\beta_1}) = 
\left(\frac{\partial r^2(y, x, \hat{\beta_0}, \hat{\beta_1})}{\partial \beta_0} \quad , \quad  
\frac{\partial r^2(y, x, \hat{\beta_0}, \hat{\beta_1})}{\partial \beta_1}\right) 
&&
\end{flalign*}

\visible<1>{
\medskip
Alternatively: Based on log-likelihood.
}
\visible<2>{
\vspace{-1cm}
$$\hspace{-0.3cm}{\color{red}\Downarrow} \hspace{3.4cm} {\color{red}\Downarrow} $$
$$
{\color{red} -2 \cdot r(y,x,\hat{\beta_0},\hat{\beta_1})}  \hspace{0.7cm} {\color{red} -2 \cdot x \cdot r(y,x,\hat{\beta_0},\hat{\beta_1})}
$$
}
\end{frame}

@@

\only<2>{
\small
$$
s(y, x, \hat{\beta_0}, \hat{\beta_1}) = 
-2 \cdot \begin{pmatrix} 
r(y_1, x_1, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} x \cdot r(y_1, x_1, \hat{\beta_0}, \hat{\beta_1})}\\
r(y_2, x_2, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} x \cdot r(y_2, x_2, \hat{\beta_0}, \hat{\beta_1})}\\
\vdots & {\color{lightgray} \vdots} \\
r(y_n, x_n, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} x \cdot r(y_n, x_n, \hat{\beta_0}, \hat{\beta_1})}
\end{pmatrix}
%\begin{pmatrix}
%\partial_{\beta_0} r^2(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} \partial_{\beta_1} r^2(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1})}\\
%\partial_{\beta_0} r^2(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} \partial_{\beta_1} r^2(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1})}\\
%\vdots & {\color{lightgray} \vdots} \\
%\partial_{\beta_0} r^2(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} \partial_{\beta_1} r^2(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1})}
%\end{pmatrix}
$$
}
\only<3>{
\small
$$
s(y, x, \hat{\beta_0}, \hat{\beta_1}) = 
-2 \cdot \begin{pmatrix} 
r(y_1, x_1, \hat{\beta_0}, \hat{\beta_1}) & x \cdot r(y_1, x_1, \hat{\beta_0}, \hat{\beta_1})\\
r(y_2, x_2, \hat{\beta_0}, \hat{\beta_1}) & x \cdot r(y_2, x_2, \hat{\beta_0}, \hat{\beta_1})\\
\vdots & \vdots \\
r(y_n, x_n, \hat{\beta_0}, \hat{\beta_1}) & x \cdot r(y_n, x_n, \hat{\beta_0}, \hat{\beta_1})
\end{pmatrix}
%\begin{pmatrix} 
%\partial_{\beta_0} r^2(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) & \partial_{\beta_1} r^2(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1})\\
%\partial_{\beta_0} r^2(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) & \partial_{\beta_1} r^2(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1})\\
%\vdots & \vdots \\
%\partial_{\beta_0} r^2(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) & \partial_{\beta_1} r^2(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1})
%\end{pmatrix}
$$
}
\only<4-5>{
\small
$$
r(y, x, \hat{\beta_0}, \hat{\beta_1}) = 
\begin{pmatrix}
r(y_1, x_1, \hat{\beta_0}, \hat{\beta_1})\\
r(y_2, x_2, \hat{\beta_0}, \hat{\beta_1})\\
\vdots \\
r(y_n, x_n, \hat{\beta_0}, \hat{\beta_1})
\end{pmatrix}
%
\visible<5>{
\quad \Rightarrow \quad
\small
\begin{pmatrix}
\text{pos}\\
\text{non-pos}\\
\vdots \\
\text{pos}
\end{pmatrix}
}
$$
}
\only<6-7>{
\small
$$
z_j = 
\begin{pmatrix}
z_{j1}\\
z_{j2}\\
\vdots \\
z_{jn}\\
\end{pmatrix}
%
\visible<7>{
\quad \Rightarrow \quad
\small
\begin{pmatrix}
\text{Q}3\\
\text{Q}1\\
\vdots \\
\text{Q}2
\end{pmatrix}
}
$$
}
\end{frame}

@@

\begin{figure}%[t]
\setkeys{Gin}{width=1\linewidth}
\minipage{0.36\textwidth}
<<dgp_stump1, fig=TRUE, echo=FALSE, width=4, height=3>>=
#par(mar=c(3,3,0,0), oma = c(3,3,0,0))
plot(p, 
     tp_args = list(FUN = identity, width = 15), 
     ip_args = list(fill = paltrees[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1, edge_panel = ep, terminal_panel = tp)
@
\endminipage
\minipage{0.36\textwidth}
<<dgp_stump2, fig=TRUE, echo=FALSE, width=4, height=3>>=
par(mar = c(4,4,2,4))
plot(x = d$x[d$z1<=xi], y = d$y[d$z1<=xi], ylim = c(-5,5), ylab = "", xlab = "", 
     main = expression("varying"~beta[0]),
     pch = 19, col = alpha(pal_dgp[1], 0.4), cex = 0.7)
mtext("y", side = 2, line = 3)
mtext("x", side = 1, line = 2.5)
abline(coef = c(-1,1), col = pal_dgp[1], lwd = 2)
abline(coef = c(1,1), col = pal_dgp[2], lwd = 2)
points(x = d$x[d$z1>xi], y = d$y[d$z1>xi], ylim = c(-5,5), pch = 19, col = alpha(pal_dgp[2], 0.4), cex = 0.7)
legend("top", c(expression(z[1]<=xi), expression(z[1]>xi)), col = c(pal_dgp[1], pal_dgp[2]), pch = c(19), bty = "n", cex = 0.9)
mtext(expression(beta[0]==+delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = 3.2)
mtext(expression(beta[1]== 1), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = 2.2)
mtext(expression(beta[0]==-delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 0.2)
mtext(expression(beta[1]== 1), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = -0.8)
@
\endminipage

@@

par(mar = c(4,4,2,4))
plot(x = d$x[d$z1<=xi], y = d$y[d$z1<=xi], ylim = c(-5,5), ylab = "", xlab = "", main = expression("varying"~beta[1]),
     pch = 19, col = alpha(pal_dgp[1], 0.4), cex = 0.7)
mtext("y", side = 2, line = 3)
mtext("x", side = 1, line = 2.5)
abline(coef = c(0,1), col = pal_dgp[1], lwd = 2)
abline(coef = c(0,-1), col = pal_dgp[2], lwd = 2)
points(x = d$x[d$z1>xi], y = d$y[d$z1>xi], ylim = c(-5,5), pch = 19, col = alpha(pal_dgp[2], 0.4), cex = 0.7)
legend("top", c(expression(z[1]<=xi), expression(z[1]>xi)), col = c(pal_dgp[1], pal_dgp[2]), pch = c(19), bty = "n", cex = 0.9)
mtext(expression(beta[0]==0), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -1)
mtext(expression(beta[1]==-delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -2)
mtext(expression(beta[0]==0), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 2)
mtext(expression(beta[1]==+delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 1)
@
\endminipage
\minipage{0.36\textwidth}
<<dgp_stump4, fig=TRUE, echo=FALSE, width=4, height=3>>=
d <- dgp_stump(nobs = nobs, binary_beta = TRUE, binary_regressor = FALSE, only_intercept = FALSE,
               vary_beta = "all", delta = delta, xi = xi, sigma = sigma)
par(mar = c(4,4,2,4))
plot(x = d$x[d$z1<=xi], y = d$y[d$z1<=xi], ylim = c(-5,5), ylab = "", xlab = "",
     main = expression("varying"~beta[0]~"and"~beta[1]),
     xlim = c(-1,1.1),
     pch = 19, col = alpha(pal_dgp[1], 0.4), cex = 0.7)
mtext("y", side = 2, line = 3)
mtext("x", side = 1, line = 2.5)
abline(coef = c(-1,1), col = pal_dgp[1], lwd = 2)
abline(coef = c(1,-1), col = pal_dgp[2], lwd = 2)
points(x = d$x[d$z1>xi], y = d$y[d$z1>=xi], pch = 19, col = alpha(pal_dgp[2], 0.4), cex = 0.7)
legend("top", c(expression(z[1]<=xi), expression(z[1]>xi)), col = c(pal_dgp[1], pal_dgp[2]), pch = c(19), bty = "n", cex = 0.9)
mtext(expression(beta[0]==-delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 1.7)
mtext(expression(beta[1]==+delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 0.7)
mtext(expression(beta[0]==+delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -0.7)
mtext(expression(beta[1]==-delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -1.7)
@
\endminipage
\end{figure}
\end{frame}

@@

\begin{figure}%[t]
\begin{center}
\setkeys{Gin}{width=1\linewidth}
\minipage{0.45\textwidth}
<<dgp_tree1, fig=TRUE, echo=FALSE, width=5, height=4.5>>=
par(mar=c(0,0,0,0), oma=c(0,0,0.4,0))
#paltrees <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))
paltrees <- rgb(c(1, 0.9, 1), c(1, 0.9, 1), c(1, 0.9, 1))
plot(p, edge_panel = ep, terminal_panel = tp,
     tp_args = list(FUN = identity, width = 15), 
     ip_args = list(fill = paltrees[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1)
@
\endminipage
\minipage{0.4\textwidth}
<<dgp_tree2, fig=TRUE, echo=FALSE, width=4, height=4>>=
par(mar = c(4,4,2,4))
plot(x = d$x[d$z2<=xi], y = d$y[d$z2<=xi], ylim = c(-5,5), ylab = "y", xlab = "x",
     pch = 19, col = alpha(pal_dgp[1], 0.4), cex = 0.7)
abline(coef = c(0,1), col = pal_dgp[1], lwd = 2)
abline(coef = c(-1,-1), col = pal_dgp[2], lwd = 2)
abline(coef = c(1,-1), col = pal_dgp[3], lwd = 2)
points(x = d$x[d$z2>xi & d$z1<=xi], y = d$y[d$z2>xi & d$z1<=xi], pch = 19, 
       col = alpha(pal_dgp[2], 0.4), cex = 0.7)
points(x = d$x[d$z2>xi & d$z1>xi], y = d$y[d$z2>xi & d$z1>xi], pch = 19, 
       col = alpha(pal_dgp[3], 0.4), cex = 0.7)
legend("top", c(expression(z[2]<=xi), 
                expression(z[2]>xi~" and "~z[1]<=xi),
                expression(z[2]>xi~" and "~z[1]>xi)), 
       col = c(pal_dgp[1], pal_dgp[2], pal_dgp[3]), pch = c(19), bty = "n", cex = 0.9)
mtext(expression(beta[0]==0), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 2)
mtext(expression(beta[1]==+delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 1.3)
mtext(expression(beta[0]==+delta), side = 4, col = pal_dgp[3], las = 1, line = 0.5, at = 0.1)
mtext(expression(beta[1]==-delta), side = 4, col = pal_dgp[3], las = 1, line = 0.5, at = -0.6)
mtext(expression(beta[0]==-delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -2)
mtext(expression(beta[1]==-delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -2.7)
@
\endminipage
\end{center}
\end{figure}

@@

Based on the simulation results the following general conclusions can be drawn:
\begin{itemize}
\item Including full model scores is advantageous over using only residuals.
\item Binarizing scores/residuals does not lead to any improvements in performance.
\item Categorizing split variables can facilitate calculations, however, important
effects might be missed unless a binning point is close to the true split point.
\item If the applied statistical test performs well as significance test pre-pruning
is preferable. Otherwise post-pruning can compensate this deficiency at least partly.
\end{itemize}
\end{frame}

>>>>

{\bf General testing strategy:}
\begin{enumerate}
\item Evaluate a discrepancy measure giving information on the goodness of fit of the model for each observation and each model parameter (transformation~of~$Y$).
\item Apply a statistical test assessing dependency of the discrepancy measure to each possible split variable $Z_j$.
\item Select the split variable $Z_j^\ast$ showing the smallest $p$-value.
\end{enumerate}

@@

{\bf Residuals:}
\begin{flalign*}
\hspace{0.7cm}
r(Y,X,\hat{\beta_0},\hat{\beta_1}) = Y - \hat{\beta_0} - \hat{\beta_1} \cdot X &&
\end{flalign*}

@@

{\bf Model scores} (based on least squares):
\begin{flalign*}
\hspace{0.7cm}
s(Y, X, \hat{\beta_0}, \hat{\beta_1}) = 
\left(\frac{\partial r^2(Y, X, \hat{\beta_0}, \hat{\beta_1})}{\partial \beta_0} \quad , \quad  
\frac{\partial r^2(Y, X, \hat{\beta_0}, \hat{\beta_1})}{\partial \beta_1}\right) 
&&
\end{flalign*}

\visible<1>{
\medskip
Alternatively: Based on log-likelihood.
}
\visible<2>{
\vspace{-1cm}
$$\hspace{-0.3cm}{\color{red}\Downarrow} \hspace{3.4cm} {\color{red}\Downarrow} $$
$$
{\color{red} -2 \cdot r(Y, X, \hat{\beta_0},\hat{\beta_1})}  \hspace{0.7cm} {\color{red} -2 \cdot X \cdot r(Y, X, \hat{\beta_0},\hat{\beta_1})}
$$
}
\end{frame}

@@

\only<2>{
\small
$$
s(Y, X, \hat{\beta_0}, \hat{\beta_1}) = 
-2 \cdot \begin{pmatrix} 
r(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} X \cdot r(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1})}\\
r(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} X \cdot r(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1})}\\
\vdots & {\color{lightgray} \vdots} \\
r(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} X \cdot r(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1})}
\end{pmatrix}
%\begin{pmatrix}
%\partial_{\beta_0} r^2(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} \partial_{\beta_1} r^2(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1})}\\
%\partial_{\beta_0} r^2(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} \partial_{\beta_1} r^2(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1})}\\
%\vdots & {\color{lightgray} \vdots} \\
%\partial_{\beta_0} r^2(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} \partial_{\beta_1} r^2(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1})}
%\end{pmatrix}
$$
}
\only<3>{
\small
$$
s(Y, X, \hat{\beta_0}, \hat{\beta_1}) = 
-2 \cdot \begin{pmatrix} 
r(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) & X \cdot r(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1})\\
r(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) & X \cdot r(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1})\\
\vdots & \vdots \\
r(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) & X \cdot r(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1})
\end{pmatrix}
%\begin{pmatrix} 
%\partial_{\beta_0} r^2(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) & \partial_{\beta_1} r^2(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1})\\
%\partial_{\beta_0} r^2(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) & \partial_{\beta_1} r^2(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1})\\
%\vdots & \vdots \\
%\partial_{\beta_0} r^2(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) & \partial_{\beta_1} r^2(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1})
%\end{pmatrix}
$$
}
\only<4-5>{
\small
$$
r(Y, X, \hat{\beta_0}, \hat{\beta_1}) = 
\begin{pmatrix}
r(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1})\\
r(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1})\\
\vdots \\
r(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1})
\end{pmatrix}
%
\visible<5>{
\quad \Rightarrow \quad
\small
\begin{pmatrix}
\text{pos}\\
\text{non-pos}\\
\vdots \\
\text{pos}
\end{pmatrix}
}
$$
}
\only<6-7>{
\small
$$
Z_j = 
\begin{pmatrix}
Z_{j1}\\
Z_{j2}\\
\vdots \\
Z_{jn}\\
\end{pmatrix}
%
\visible<7>{
\quad \Rightarrow \quad
\small
\begin{pmatrix}
\text{Q}3\\
\text{Q}1\\
\vdots \\
\text{Q}2
\end{pmatrix}
}
$$
}
\end{frame}

@@

\begin{figure}%[t]
\setkeys{Gin}{width=1\linewidth}
\minipage{0.36\textwidth}
<<dgp_stump1, fig=TRUE, echo=FALSE, width=4, height=3>>=
#par(mar=c(3,3,0,0), oma = c(3,3,0,0))
plot(p, 
     tp_args = list(FUN = identity, width = 15), 
     ip_args = list(fill = paltrees[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1, edge_panel = ep, terminal_panel = tp)
@
\endminipage
\minipage{0.36\textwidth}
<<dgp_stump2, fig=TRUE, echo=FALSE, width=4, height=3>>=
par(mar = c(4,4,2,4))
plot(x = d$x[d$z1<=xi], y = d$y[d$z1<=xi], ylim = c(-5,5), ylab = "", xlab = "", 
     main = expression("varying"~beta[0]),
     pch = 19, col = alpha(pal_dgp[1], 0.4), cex = 0.7)
mtext("Y", side = 2, line = 2.5)
mtext("X", side = 1, line = 2.5)
abline(coef = c(-1,1), col = pal_dgp[1], lwd = 2)
abline(coef = c(1,1), col = pal_dgp[2], lwd = 2)
points(x = d$x[d$z1>xi], y = d$y[d$z1>xi], ylim = c(-5,5), pch = 19, col = alpha(pal_dgp[2], 0.4), cex = 0.7)
legend("top", c(expression(z[1]<=xi), expression(z[1]>xi)), col = c(pal_dgp[1], pal_dgp[2]), pch = c(19), bty = "n", cex = 0.9)
mtext(expression(beta[0]==+delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = 3.2)
mtext(expression(beta[1]== 1), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = 2.2)
mtext(expression(beta[0]==-delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 0.2)
mtext(expression(beta[1]== 1), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = -0.8)
@
\endminipage

@@

par(mar = c(4,4,2,4))
plot(x = d$x[d$z1<=xi], y = d$y[d$z1<=xi], ylim = c(-5,5), ylab = "", xlab = "", main = expression("varying"~beta[1]),
     pch = 19, col = alpha(pal_dgp[1], 0.4), cex = 0.7)
mtext("Y", side = 2, line = 2.5)
mtext("X", side = 1, line = 2.5)
abline(coef = c(0,1), col = pal_dgp[1], lwd = 2)
abline(coef = c(0,-1), col = pal_dgp[2], lwd = 2)
points(x = d$x[d$z1>xi], y = d$y[d$z1>xi], ylim = c(-5,5), pch = 19, col = alpha(pal_dgp[2], 0.4), cex = 0.7)
legend("top", c(expression(z[1]<=xi), expression(z[1]>xi)), col = c(pal_dgp[1], pal_dgp[2]), pch = c(19), bty = "n", cex = 0.9)
mtext(expression(beta[0]==0), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -1)
mtext(expression(beta[1]==-delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -2)
mtext(expression(beta[0]==0), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 2)
mtext(expression(beta[1]==+delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 1)
@
\endminipage
\minipage{0.36\textwidth}
<<dgp_stump4, fig=TRUE, echo=FALSE, width=4, height=3>>=
d <- dgp_stump(nobs = nobs, binary_beta = TRUE, binary_regressor = FALSE, only_intercept = FALSE,
               vary_beta = "all", delta = delta, xi = xi, sigma = sigma)
par(mar = c(4,4,2,4))
plot(x = d$x[d$z1<=xi], y = d$y[d$z1<=xi], ylim = c(-5,5), ylab = "", xlab = "",
     main = expression("varying"~beta[0]~"and"~beta[1]),
     xlim = c(-1,1.1),
     pch = 19, col = alpha(pal_dgp[1], 0.4), cex = 0.7)
mtext("Y", side = 2, line = 2.5)
mtext("X", side = 1, line = 2.5)
abline(coef = c(-1,1), col = pal_dgp[1], lwd = 2)
abline(coef = c(1,-1), col = pal_dgp[2], lwd = 2)
points(x = d$x[d$z1>xi], y = d$y[d$z1>=xi], pch = 19, col = alpha(pal_dgp[2], 0.4), cex = 0.7)
legend("top", c(expression(z[1]<=xi), expression(z[1]>xi)), col = c(pal_dgp[1], pal_dgp[2]), pch = c(19), bty = "n", cex = 0.9)
mtext(expression(beta[0]==-delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 1.7)
mtext(expression(beta[1]==+delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 0.7)
mtext(expression(beta[0]==+delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -0.7)
mtext(expression(beta[1]==-delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -1.7)
@
\endminipage
\end{figure}
\end{frame}

@@

\begin{figure}%[t]
\begin{center}
\setkeys{Gin}{width=1\linewidth}
\minipage{0.45\textwidth}
<<dgp_tree1, fig=TRUE, echo=FALSE, width=5, height=4.5>>=
par(mar=c(0,0,0,0), oma=c(0,0,0.4,0))
#paltrees <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))
paltrees <- rgb(c(1, 0.9, 1), c(1, 0.9, 1), c(1, 0.9, 1))
plot(p, edge_panel = ep, terminal_panel = tp,
     tp_args = list(FUN = identity, width = 15), 
     ip_args = list(fill = paltrees[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1)
@
\endminipage
\minipage{0.4\textwidth}
<<dgp_tree2, fig=TRUE, echo=FALSE, width=4, height=4>>=
par(mar = c(4,4,2,4))
plot(x = d$x[d$z2<=xi], y = d$y[d$z2<=xi], ylim = c(-5,5), ylab = "Y", xlab = "X",
     pch = 19, col = alpha(pal_dgp[1], 0.4), cex = 0.7)
abline(coef = c(0,1), col = pal_dgp[1], lwd = 2)
abline(coef = c(-1,-1), col = pal_dgp[2], lwd = 2)
abline(coef = c(1,-1), col = pal_dgp[3], lwd = 2)
points(x = d$x[d$z2>xi & d$z1<=xi], y = d$y[d$z2>xi & d$z1<=xi], pch = 19, 
       col = alpha(pal_dgp[2], 0.4), cex = 0.7)
points(x = d$x[d$z2>xi & d$z1>xi], y = d$y[d$z2>xi & d$z1>xi], pch = 19, 
       col = alpha(pal_dgp[3], 0.4), cex = 0.7)
legend("top", c(expression(z[2]<=xi), 
                expression(z[2]>xi~" and "~z[1]<=xi),
                expression(z[2]>xi~" and "~z[1]>xi)), 
       col = c(pal_dgp[1], pal_dgp[2], pal_dgp[3]), pch = c(19), bty = "n", cex = 0.9)
mtext(expression(beta[0]==0), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 2)
mtext(expression(beta[1]==+delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 1.3)
mtext(expression(beta[0]==+delta), side = 4, col = pal_dgp[3], las = 1, line = 0.5, at = 0.1)
mtext(expression(beta[1]==-delta), side = 4, col = pal_dgp[3], las = 1, line = 0.5, at = -0.6)
mtext(expression(beta[0]==-delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -2)
mtext(expression(beta[1]==-delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -2.7)
@
\endminipage
\end{center}
\end{figure}

@@

Based on the simulation results the following general conclusions can be drawn:
\begin{itemize}
\item Including full model scores is advantageous over using only residuals.
\item Binarizing scores/residuals does not lead to any improvements in performance.
\item Categorizing split variables can simplify calculations, however, important
effects might be missed unless a binning point is close to the true split point.
\item If the applied statistical test performs well as significance test pre-pruning
is preferable. Otherwise post-pruning can compensate this deficiency at least partly.
\end{itemize}
\end{frame}
